\chapter{Homomorphic Authenticators}\label{chap:homo-auth}

A trivial solution to the problem of enforcing authenticity of outsourced
computations would be to make the server send the entire data and the computed
signatures/MACs (we will refer to both of them as tags from now on) back to the
client, and then she would compute the function herself and compare the
results. But by using homomorphic authentication primitives, and by sending
only the tags and the output of the function back to the client, beyond saving
bandwidth, the amount of information revealed to the client about the data set
is much smaller. Also, this trivial solution does not guarantee efficiency
which is one of the requirements to enforce the authenticity of computations.

% Remove this paragraph if using the Definitions section
Let us now introduce some basic notions and definitions necessary to build
homomorphic authentication schemes, as well as some constructions. From now on,
it is always assumed that Alice stores the data and tags on some possibly
untrusted server.

\input{Chapters/_homo-sigs}
\input{Chapters/_homo-macs}

%% ---------------------------------------------------------------------------
% LIMITATIONS AND OPEN PROBLEMS
%% ---------------------------------------------------------------------------
\section{Limitations and open problems}\label{sec:limitations}
Homomorphic signatures are still far from being a truly practical primitive.
The data set size has to be fixed a-priori, and besides the
\citescheme{boneh:freeman:2011} scheme with support for polynomials of bounded
degree, only linear realizations have been proposed. However, for quadratic and
higher degree polynomials, the derived signatures leak information about the
original data set. This privacy property is only achieved for linear functions,
and therefore, it remains an open problem to extend it for quadratic and higher
polynomials.

The security of \citescheme{boneh:freeman:2011} scheme could be strengthened by
removing the random oracle model which is used to simulate signatures on
a chosen message attacker, but it is not known if that's even possible without
losing the homomorphic properties.

In the private key setting, the existence of a fully homomorphic MAC scheme is
already known as it was shown by \textcite{gennaro:wichs:2012}. However, its
security model does not allow an adversary to ask for verification queries, and
it would be interesting to improve the \citescheme{gennaro:wichs:2012} scheme
to support an unbounded number of verification queries. Just as with
the \citescheme{catalano:fiore:2013} constructions, the complexity of the
verification algorithm of a fully homomorphic MAC is not independent of the
complexity of a program $\calP$. This, and the fact that it relies on FHE
``machinery'', makes the \citescheme{gennaro:wichs:2012} scheme not practical.

The~\citescheme{catalano:fiore:2013}
and~\citescheme{backes:fiore:reischuk:2013} schemes only support a limited
class of functions. Even though they allow the computation of many interesting
functions, and because of their superior efficiency compared to fully
homomorphic MACs, it would be very interesting to extend them to a broader
class of functions, i.e., to construct a fully homomorphic MAC for circuits of
arbitrary size. 

Another interesting open problem is the existence of a \emph{Fully Homomorphic
Signature} scheme which could authenticate the computation of any function on
signed data. This scheme would be particularly useful to be used alongside
a FHE scheme like the one from \textcite{gentry:2009:FHE}. It is also not clear
whether processes from~\citeauthor{gentry:2009:FHE}'s construction could be
applied to signature schemes in order to improve their security and efficiency.
While that is not possible, it would be useful to enlarge the set of admissible
functions of homomorphic signatures schemes.
